Making Sequential Programs Deterministic

### 2019-04-05 java-design.tex
Intro text about implementation

We have implemented our type system for Java, in a tool named \theDeterminismChecker.

@@

In Java, a type qualifier is written as a type annotation (which has a leading \<@>), as in
\<@Positive int>.

Our implementation
handles array types, and any class (including user-defined classes) that
implements the \<Iterable> or \<Iterator> interfaces; this includes all
Java collections such as \<List>s and \<Set>s.  For simplicity, and because
the semantics and rules are similar even though the syntax varies, this
paper uses the term ``collection'' and the type \<Collection> to represent
all these types.

(In Java,
\<HashMap> cannot be deterministic, but \<TreeMap> and \<LinkedHashMap>
can.\todo{forward reference?})

>>>>

We have implemented our type system for Java, in a tool named \theDeterminismChecker.
The implementation consists of 1200 lines of Java
code, plus 1800 lines of tests (all line measurements are non-comment,
non-blank lines), a manual, examples, etc.  It is publicly available at
\myurl{URL for \theDeterminismChecker}.

@@

\subsection{Determinism type qualifiers}

A type qualifier is written in Java source code as a type annotation.
A type annotation has a leading ``\<@>'' and is written immediately before a
Java basetype, as in \<@Positive int> or \codeid{@NonEmpty List<@NonNull String>}.

\TheDeterminismChecker supports the type qualifiers \<@NonDet>,
\<@OrderNonDet>, and \<@Det> (and others described below).

Our implementation treats arrays analogously to collections:
the semantics and rules are similar even though the syntax varies.
For simplicity this
paper uses the term ``collection'' and the type \<Collection> to represent
collections and arrays.
Collections include any class (including user-defined classes) that
implements the \<Iterable> or \<Iterator> interfaces; this includes all
Java collections such as \<List>s and \<Set>s.
